Revealing Transient Strain in Geodetic Data with the Radial Basis Function Finite Difference Method

### 2017-05-29 HinesHetlandGPR.tex
added more

Let $\bm{y}$ be the set of $(x_i, t)$ pairs describing where and when each of the $n$ GNSS observations have been made. Furthermore, let $\bm{a}$ be the vector of coefficients from eq. (\ref{eq:Data}) for each of the $m$ GNSS monuments, and let $\bm{P}$ be the $n \times 6m$ matrix of corresponding basis functions evaluated at each point in $\bm{y}$. The $n$ GNSS observations, $\bm{d}_*$, are a realization of the random vector
\begin{equation}
\bm{d} = u(\bm{y}) + \epsilon(\bm{y}) + \bm{P}\bm{a},
\end{equation}
which is distributed as $\mathcal{N}(\bm{0},\bm{C_d})$, where
\begin{equation}\label{eq:Cd}
\bm{C_d} = C_u(\bm{y},\bm{y}) + C_\epsilon(\bm{y},\bm{y}) + 
           \lambda^2\bm{P}\bm{P}^T.
\end{equation}
It should be understood that the notation $f(\bm{x})$ and $f(\bm{x},\bm{y})$ is used as shorthand for $[f(x_i)]_{x_i \in \bm{x}}$ and $[f(x_i,y_j)]_{(x_i,y_j) \in \bm{x} \times \bm{y}}$, respectively. In section X, we describe how the observations $\bm{d}_*$ are used to constrain transient displacements $u$ and then for estimates of transient strain rates.

@@

\subsection{Conditioning}
In this section we describe how to condition the prior for transient displacements, $u$, with GNSS observations, $\bm{d}_*$, to form a posterior estimate of transient displacements, $\hat{u} = u | \bm{d}_*$. We will assume that an appropriate covariance functions and corresponding hyperparameters for $X$, $T$, and $C_\epsilon$ have already been chosen. If $\lambda$ is kept finite, then we can follow \citet{Rasmussen2006} to find that $\hat{u}$ is distributed as $\mathcal{N}(\mu_{\hat{u}},C_{\hat{u}})$, where
\begin{equation}\label{eq:PosteriorMean}
\mu_{\hat{u}}(z) = C_u(z,\bm{y})\bm{C_d}^{-1}\bm{d}_*
\end{equation}    
and
\begin{equation}\label{eq:PosteriorCov}
C_{\hat{u}}(z,z') = C_u(z,\bm{y})\bm{C_d}^{-1}C_u(\bm{y},z').
\end{equation}
However, we are interested in the limit as $\lambda \to \infty$, and the form for eq. (\ref{eq:PosteriorMean}) and eq. (\ref{eq:PosteriorCov}) is not suitable for evaluating this limit. We can use the partitioned matrix inversion identity \citep[e.g.,][]{Press2007} to rewrite eq. (\ref{eq:PosteriorMean}) and eq. (\ref{eq:PosteriorCov}) as
 \begin{equation}\label{eq:PosteriorMean2}
\mu_{\hat{u}}(z) =
\left[ 
\begin{array}{cc}
C_u(z,\bm{y}) & \bm{0} \\
\end{array}
\right]
\left[
\begin{array}{cc}
C_u(\bm{y},\bm{y}) + C_\epsilon(\bm{y},\bm{y}) & \bm{P} \\
\bm{P}^T  & -\lambda^{-2} \bm{I} \\
\end{array}
\right]^{-1}
\left[
\begin{array}{c}
\bm{d}_* \\
\bm{0} \\
\end{array}
\right]
\end{equation}    
and
\begin{equation}\label{eq:PosteriorCov2}
C_{\hat{u}}(z,z') = 
C_u(z,z') - 
\left[ 
\begin{array}{cc}
C_u(z,\bm{y}) & \bm{0} \\
\end{array}
\right]
\left[
\begin{array}{cc}
C_u(\bm{y},\bm{y}) + C_\epsilon(\bm{y},\bm{y}) & \bm{P} \\
\bm{P}^T  & -\lambda^{-2} \bm{I} \\
\end{array}
\right]^{-1}
\left[
\begin{array}{c}
C_u(\bm{y},z') \\
\bm{0} \\
\end{array}
\right].
\end{equation}
Taking the limit as $\lambda \to \infty$, we get the solution for the mean and covariance of $\hat{u}$,
 \begin{equation}\label{eq:PosteriorMean3}
\mu_{\hat{u}}(z) =
\left[ 
\begin{array}{cc}
C_u(z,\bm{y}) & \bm{0} \\
\end{array}
\right]
\left[
\begin{array}{cc}
C_u(\bm{y},\bm{y}) + C_\epsilon(\bm{y},\bm{y}) & \bm{P} \\
\bm{P}^T  & \bm{0} \\
\end{array}
\right]^{-1}
\left[
\begin{array}{c}
\bm{d}_* \\
\bm{0} \\
\end{array}
\right]
\end{equation}    
and
\begin{equation}\label{eq:PosteriorCov3}
C_{\hat{u}}(z,z') = 
C_u(z,z') - 
\left[ 
\begin{array}{cc}
C_u(z,\bm{y}) & \bm{0} \\
\end{array}
\right]
\left[
\begin{array}{cc}
C_u(\bm{y},\bm{y}) + C_\epsilon(\bm{y},\bm{y}) & \bm{P} \\
\bm{P}^T  & \bm{0} \\
\end{array}
\right]^{-1}
\left[
\begin{array}{c}
C_u(\bm{y},z') \\
\bm{0} \\
\end{array}
\right].
\end{equation}

which allows us to evaluate the limit  evaluate the limit 
and we consider the GNSS observations $\bm{d}_*$ to be a realization of $\bm{d}$. Furthermore, we The random vector $\bm{d}$ is distributed as $\bm{d} \sim \mathcal{N}(\bar{\bm{d}},\bm{C_d})$, where 
\begin{equation}
\bm{d}
\end{equation}

We consider the column vector of GNSS observations, $\bm{d}_*$, to be realizations of $d_{ij}$ at the positions and times in $\bm{y}$ (i.e, $\bm{d}_* = [d_{ij}(\omega_*)]^T_{(x_i,t_j) \in \bm{y}}$). We want to find the posterior estimate of transient displacements, $u | \bm{d}_*$, which incorporates our prior and the GNSS observations.

Once we have chosen covariance functions and corresponding hyperparameters for $X$, $T$, and $C_d$, we then condition our prior Gaussian process, $u$, with the observations $d_*$ to form a posterior estimate of transient displacements, $u|d_*$. $d_*$ is a realization of the

Whichever the Gaussian processes describing transient displacements, $u$, and noise in the GNSS data, $\epsilon$, have covariance functions which

Optimal Hyperparameters

Conditioning the GP

Differentiation, scaling and addition to get strain

>>>>

Let $\bm{y}$ be the set of $(x_i, t)$ pairs describing where and when each of the $n$ GNSS observations have been made. Furthermore, let $\bm{a}$ be the vector of coefficients from eq. (\ref{eq:Data}) for each of the $m$ GNSS monuments, and let $\bm{P}$ be the $n \times 6m$ matrix of corresponding basis functions evaluated at each point in $\bm{y}$. The $n$ GNSS observations, $\bm{d}_*$, are a realization of the random vector
\begin{equation}
\bm{d} = u(\bm{y}) + \epsilon(\bm{y}) + \bm{P}\bm{a},
\end{equation}
which is distributed as $\mathcal{N}(\bm{0},\bm{C_d})$, where
\begin{equation}\label{eq:Cd}
\bm{C_d} = C_u(\bm{y},\bm{y}) + C_\epsilon(\bm{y},\bm{y}) + 
           \lambda^2\bm{P}\bm{P}^T.
\end{equation}
It should be understood that the notation $f(\bm{x})$ and $f(\bm{x},\bm{y})$ is shorthand for $[f(x_i)]_{x_i \in \bm{x}}$ and $[f(x_i,y_j)]_{(x_i,y_j) \in \bm{x} \times \bm{y}}$, respectively. In section X, we describe how the observations $\bm{d}_*$ are used to constrain transient displacements $u$ and then for estimates of transient strain rates.

We now describe how to condition the prior for transient displacements, $u$, with GNSS observations, $\bm{d}_*$, to form a posterior estimate of transient displacements, $\hat{u} = u | \bm{d}_*$. We will assume that an appropriate covariance functions and corresponding hyperparameters for $X$, $T$, and $C_\epsilon$ have already been chosen. If $\lambda$ is kept finite, then we can follow \citet{Rasmussen2006} to find that $\hat{u}$ is distributed as $\mathcal{N}(\mu_{\hat{u}},C_{\hat{u}})$, where
\begin{equation}\label{eq:PosteriorMean}
\mu_{\hat{u}}(z) = C_u(z,\bm{y})\bm{C_d}^{-1}\bm{d}_*
\end{equation}    
and
\begin{equation}\label{eq:PosteriorCov}
C_{\hat{u}}(z,z') = C_u(z,\bm{y})\bm{C_d}^{-1}C_u(\bm{y},z').
\end{equation}
However, we are interested in the limit as $\lambda \to \infty$, and the form for eq. (\ref{eq:PosteriorMean}) and eq. (\ref{eq:PosteriorCov}) is not suitable for evaluating this limit. We can use the partitioned matrix inversion identity \citep[e.g.,][]{Press2007} to rewrite eq. (\ref{eq:PosteriorMean}) and eq. (\ref{eq:PosteriorCov}) as
 \begin{equation}\label{eq:PosteriorMean2}
\mu_{\hat{u}}(z) =
\left[ 
\begin{array}{cc}
C_u(z,\bm{y}) & \bm{0} \\
\end{array}
\right]
\left[
\begin{array}{cc}
C_u(\bm{y},\bm{y}) + C_\epsilon(\bm{y},\bm{y}) & \bm{P} \\
\bm{P}^T  & -\lambda^{-2} \bm{I} \\
\end{array}
\right]^{-1}
\left[
\begin{array}{c}
\bm{d}_* \\
\bm{0} \\
\end{array}
\right]
\end{equation}    
and
\begin{equation}\label{eq:PosteriorCov2}
C_{\hat{u}}(z,z') = 
C_u(z,z') - 
\left[ 
\begin{array}{cc}
C_u(z,\bm{y}) & \bm{0} \\
\end{array}
\right]
\left[
\begin{array}{cc}
C_u(\bm{y},\bm{y}) + C_\epsilon(\bm{y},\bm{y}) & \bm{P} \\
\bm{P}^T  & -\lambda^{-2} \bm{I} \\
\end{array}
\right]^{-1}
\left[
\begin{array}{c}
C_u(\bm{y},z') \\
\bm{0} \\
\end{array}
\right].
\end{equation}
Taking the limit as $\lambda \to \infty$, we get the solution for the mean and covariance of $\hat{u}$,
 \begin{equation}\label{eq:PosteriorMean3}
\mu_{\hat{u}}(z) =
\left[ 
\begin{array}{cc}
C_u(z,\bm{y}) & \bm{0} \\
\end{array}
\right]
\left[
\begin{array}{cc}
C_u(\bm{y},\bm{y}) + C_\epsilon(\bm{y},\bm{y}) & \bm{P} \\
\bm{P}^T  & \bm{0} \\
\end{array}
\right]^{-1}
\left[
\begin{array}{c}
\bm{d}_* \\
\bm{0} \\
\end{array}
\right]
\end{equation}    
and
\begin{equation}\label{eq:PosteriorCov3}
C_{\hat{u}}(z,z') = 
C_u(z,z') - 
\left[ 
\begin{array}{cc}
C_u(z,\bm{y}) & \bm{0} \\
\end{array}
\right]
\left[
\begin{array}{cc}
C_u(\bm{y},\bm{y}) + C_\epsilon(\bm{y},\bm{y}) & \bm{P} \\
\bm{P}^T  & \bm{0} \\
\end{array}
\right]^{-1}
\left[
\begin{array}{c}
C_u(\bm{y},z') \\
\bm{0} \\
\end{array}
\right].
\end{equation}

@@
